\documentclass[a4paper,11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage[backend=bibtex]{biblatex}
\bibliography{biblio}

% for probability trees
\usepackage{tikz}
\usetikzlibrary{trees}

%for strikethrough text
\usepackage{soul}

%for R source code listing
\usepackage{listings}

% For not indenting the first line of paragraphs:
\setlength{\parindent}{0pt}
% define the title
\author{John Hancock}
\title{MIT Introduction to Statistics 18.05 Reading 3 - 
\textit{Think} questions }
\begin{document}
% generates the title
\maketitle
% insert the table of contents
\tableofcontents
\section{References and License}
We are answering questions in the material from MIT OpenCourseWare
course 18.05, Introduction to Probability and Statistics.

Please see the references section for detailed citation information.

The material for the course is licensed under the terms at 
\url{http://ocw.mit.edu/terms}.

\section{\textit{Think} What is $P\left(S_{2} \mid S_{1}^{\complement}\right)$}

$S_{1}$ = 'first card is a spade'
$S_{2}$ = 'second card is a spade'

Therefore $S_{2}^{\complement}$  = 'second card is not a spade'

We will calculate

\begin{equation}
  P\left(S_{2} \mid S_{1}^{\complement}\right)
\end{equation}

That is, we will calculate the probability that the first card is a
spade given that the second card is not a spade.

We will apply the same method Orloff and Bloom use in \cite{reading3}, 
section 3, "Multiplication Rule."

If the first card is a spade then of the 51 cards remaining, 
$3 \times 13 = 39$ cards are not spades.

Therefore 
\begin{equation} \label{s1S2Condish}
  P\left(S_{2} \mid S_{2}^{\complement}\right) = \frac{39}{51}
\end{equation}

We use the multiplication rule from \cite{reading3} 
to compute the same probability.

We apply the multiplcation rule letting $A=S_{1}$ and $B=S_{2}^{\complement}$

\begin{equation} \label{multRule}
  P\left(S_{1} \cap S_{2}^{\complement}\right) =
  P\left(S_{1} \mid S_{2}^{\complement}\right) \cdot 
  P\left(S_{2}^{\complement}\right)
\end{equation}

$P\left( S_{1} \right) = \frac{13}{52}$  This is because $13$ of the 
$52$ cards are spades.

$ P\left( S_{2}^{\complement} \right) = \frac{39}{52}$ because $39$ 
cards are not spades, and we are dividing by $52$ because any of the 
$39$ out of $52$ cards are equallly likely in all possible sequences 
of two cards.


$P \left( S_{1} \cap S_{2}^{\complement} \right)
 = \frac{13}{52} \frac{39}{51}$
Because there are $13$ out of $52$ spades to draw for the first card,
and $39$ out of $51$ cards that are not spades to draw for the second
card.

We computed $P\left(S_{1} \mid S_{2}^{\complement} \right)$ in 
equation \ref{s1S2Condish}.  Now we will solve equation \ref{multRule} for 
$P\left(S_{1} \mid S_{2}^{\complement} \right)$:

\begin{equation} \label{multRule}
  P\left(S_{1} \mid S_{2}^{\complement}\right) = 
  \frac
  {
    P\left(S_{1} \cap S_{2}^{\complement}\right)
  }
  {
    P\left(S_{2}^{\complement}\right)
  } =
  \frac 
  { 
    \left( \frac{13}{52} \frac{39}{51} \right) 
  }
  {
    \frac{39}{52}
  } =
  \frac{13}{52} \frac{52}{39} \frac{39}{51} =
  \frac{13}{51}
\end{equation}

This is the same result we found in equation \ref{s1S2Condish}.

\section{\textit{Think:} For what other value(s) of P(A) is A 
independent of itself?}

Any value of $P\left(A\right)$ where $A$ will be independent of itself
will satisfy the equation
\begin{equation}
  P\left(A \right) = P\left(A \right)^{2}
\end{equation}

We find this equation in \cite{selfIndependent}.

Therefore members of the set $\left\{x \mid x=x^{2}, x\in \Re \right\}$ are
the values of $P$ we seek.

We solve for x:

\begin{equation} \label{selfSquare}
  x = x^{2}
\end{equation}

\begin{equation}
  \iff -x^{2} + x = 0
\end{equation}

\begin{equation}
  \iff -x\left(x - 1 \right) = 0
\end{equation}

\begin{equation}
  \iff x\left(x - 1 \right) = 0
\end{equation}

\begin{equation}
  \iff x \in \left\{0,1 \right\}
\end{equation}

Therefore another value for $P\left(A\right)$ that is independent of
itself is $P\left(A\right) = 1$.  Furthermore, because we solved
\ref{selfSquare} for $\left\{x \in \Re \right\}$, these are the only
values for $P\left( A \right)$ for which $A$ is independent of itself.
$\blacksquare$

\printbibliography{}
\end{document}
