\documentclass[a4paper,11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}

% for proofs  environment
\usepackage{amsthm}

% for 3d plots
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepgfplotslibrary{patchplots}

\usepackage[backend=bibtex]{biblatex}
\bibliography{reading7qub}

% for probability trees
\usepackage{tikz}
\usetikzlibrary{trees}

% for Venn diagrams
\usetikzlibrary{shapes,backgrounds}
% for plots
\usepackage{ pgfplots}
% inserted on suggestion in warning during compilation
\pgfplotsset{compat=1.9}

%for strikethrough text
\usepackage{soul}

%for R source code listing
\usepackage{listings}

%for block quotes
\usepackage{csquotes}

% For not indenting the first line of paragraphs:
\setlength{\parindent}{0pt}
% define the title
\author{John Hancock}
\title{MIT Introduction to Statistics 18.05 Reading 7B Questions}
\begin{document}
% generates the title
\maketitle
% insert the table of contents
\tableofcontents
\section{References and License}
We are answering questions in the material from MIT OpenCourseWare
course 18.05, Introduction to Probability and Statistics.

In this document we are answering questions Orloff and Bloom ask in
\cite{reading7qub}.

Please see the references section for detailed citation information.

The material for the course is licensed under the terms at
\url{http://ocw.mit.edu/terms}.

We use documentation in  \cite{logicNot} to write \LaTeX source code for this
document.

\section{Non-zero covariance}
The first question Orloff and Bloom ask is: if Cov$\left(X, Y \right)\neq 0$,
are $X$, and $Y$ independent?

In \cite{reading7b} Orloff and Bloom state that, if $X$, and $Y$ are
independent, then Cov$\left(X, Y \right)=0$.

We use symbolic logic to answer this question.  The validity of our answer
rests on Peter Williams' work in \cite{symbolicLog}

Let $p$ be the statement, "$X$, and $Y$ are independent", and $q$ be the
statement, "covariance is $0$."

Then, for this question, we know that $p$ implies $q$.

We use a result from \cite{contrapositive} that tells us that $p$ implies
$q$, if and only if $\neg q$ implies $\neg p$.

In this case $\neg q$ is the statement, "covariance $\neq 0$," and $\neg p$ is
the statement, "$X$, and $Y$ are not independent."

Therefore we know that if Cov$\left( X, Y \right) \neq 0$, then $X$, and $Y$
are not independent.

\section{Zero covariance}

For this question, Orloff and Bloom ask us a variation on the first, "Suppose
Cov$\left(X, Y\right)=0$ are $X$ and $Y$ independent?"

In \cite{reading7b} Orloff and Bloom state that if $X$ and $Y$ are independent,
then their covariance is 0.  They state this as a property of covariance.

However, in \cite{reading7b} Orloff and Bloom give an example where two
random variables are not independent, and have a covariance of $0$.

Therefore cases exist where two independent variables have a covariance of $0$,
and cases exist where two variables that are not independent have a covariance
of $0$.

Therefore the answer to this question is maybe.

\section{Joint-distributed discrete random variables}
For these problems, Orloff and Bloom give us the following joint distribution
of discrete random variables, $X$, and $Y$:


\printbibliography{}
\end{document}
