
\documentclass[a4paper,11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage[backend=bibtex]{biblatex}
\bibliography{biblio}

% for probability trees
\usepackage{tikz}
\usetikzlibrary{trees}

%for strikethrough text
\usepackage{soul}

%for R source code listing
\usepackage{listings}

% For not indenting the first line of paragraphs:
\setlength{\parindent}{0pt}
% define the title
\author{John Hancock}
\title{MIT Introduction to Statistics 18.05 Problem Set 1}
\begin{document}
% generates the title
\maketitle
% insert the table of contents
\tableofcontents
\section{Exact Formula For $P\left(B\right)$}
An element of $\omega \in \Omega$ is a sequence of birthdays.  

We gave a formal definition of $\Omega$ in

A birthday is an integer from $1$ through $365$. 
All birthdays are equally likely, and two or more people can be born on
the same day, so any $\omega$ of length $n$ is a sample with replacement
from the set $\left\{ 1, 2, 3, ..., 365 \right\}$. We will use
$365^{n}$ for the number of $\omega \in \Omega$ for a given $n$ is.

Using $365^{n}$ for the total number of $\omega \in \Omega$ means
that the order of birthdays in a given $\omega$ is important.
To see this, consider that for $n=2$ two elements of $\Omega$ are
$\left( 5, 6 \right)$ and $\left(6, 5\right)$.

Therefore $365^{n}$ is the denomenator we use when we are dividing our
countof events by the total number of events to calculate a probability.

It will be easier to calculate the probability $\bar{p}$ that for an
$\omega$ containing $n$ birthdays, none of the birthdays are the same.
Then the probability $p$ that some two birthdays in $\omega$ are the
same  will be $1-\bar{p}$.

It is easier to calculate $1-\bar{p}$ because in order to calculate $p$
directly, we have to take into account that there are $binom{n}{2}$
ways to select two birthdays in $\omega$ to be the same, or 
$binom{n}{3}$ ways to selct three birthdays in $\omega$ to be the same,
or so on.

To calculate $1-\bar{p}$ we must count the number of ways to select $n$
birthdays from a set of $365$ birthdays where no two birthdays are the
same.

Recall that we are using $365^{n}$ as the total number of
$\omega \in \Omega$, and the order of birthdays in a given $\omega${}
is important to us.

The number of samples of size $n$ of $365$ elements, where the order
is important is ${}_{365}P_{n}$. These samples of size $n$ meet the
definition of elements of $\Omega$.

Therefore the probability $hat{p}$ that some $\omega \in \Omega$ has
$n$ distinct birthdays is:

\begin{equation}
  \bar{p} = \frac {{}_{365}P_{n}} {365^{n}}
\end{equation}


\end{document}
