\documentclass[a4paper,11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}

% for proofs  environment
\usepackage{amsthm}

\usepackage[backend=bibtex]{biblatex}
\bibliography{slides6}

% for probability trees
\usepackage{tikz}
\usetikzlibrary{trees}

% for Venn diagrams
\usetikzlibrary{shapes,backgrounds}
% for plots
\usepackage{ pgfplots}
% inserted on suggestion in warning during compilation
\pgfplotsset{compat=1.9}

%for strikethrough text
\usepackage{soul}

%for R source code listing
\usepackage{listings}

%for block quotes
\usepackage{csquotes}

% For not indenting the first line of paragraphs:
\setlength{\parindent}{0pt}
% define the title
\author{John Hancock}
\title{MIT Introduction to Statistics 18.05 Reading 6A Think Questions }
\begin{document}
% generates the title
\maketitle
% insert the table of contents
\tableofcontents
\section{References and License}
We are answering questions in the material from MIT OpenCourseWare
course 18.05, Introduction to Probability and Statistics.

In this document we are answering questions Orloff and Bloom ask in
\cite{slides6}.

We use material in \cite{cubeRoot}
to write the \LaTeX code for this
document.

Please see the references section for detailed citation information.

The material for the course is licensed under the terms at
\url{http://ocw.mit.edu/terms}.



\section{Questions about X}

In this section we answer questions Orloff and Bloom ask in \cite{reading6a}
regarding a random variable $X$.

Orloff and Bloom specify that $X$ is defined on $\left[ 0, 1 \right]$, and
the pdf of $X$ is $cx^2$.

\subsection{Value of $c$}

Orloff and Bloom ask us to calculate the value of $c$.  We will use rules
and properties for integration from \cite{basicInt} in order to calculate
the value for $c$.

We know
\begin{equation}
  \int_0^1 cx^2 \, dx = 1.
\end{equation}

Therefore
\begin{equation}
  c \int_0^1 x^2 \, dx = 1.
\end{equation}

The anti-derivative of $x^2$ is $\frac{x^3}{3} +C$, so we can replace the
integral in the equation above with:

\begin{equation}
  c \left(\frac{x^3}{3} \bigg\rvert_0^1 \right)=1.
\end{equation}

We then evaluate the anti-derivative over the interval $\left[0, 1\right]$ to
obtain:

\begin{equation}
  c \left(\frac{1^3}{3} \right)=1.
\end{equation}

This implies $c=3$.

\subsection{Mean, variance, and standard deviation of $X$}
\subsubsection{Mean of $X$}
We use the definition of mean value that Orloff and Bloom give in
\cite{reading6a}.
The mean value of $X$ is
\begin{equation}
\mu = \int_0^1 x\left( 3x^2 \right)\,dx.
\end{equation}

We multiply the terms in the polynomial in the integral above to get:

\begin{equation}
\mu = \int_0^1 \left( 3x^3 \right)\,dx.
\end{equation}

We replace the integral above with its anti-derivative:

\begin{equation}
\mu = \frac{3x^4}{4} \bigg\rvert_0^1.
\end{equation}

We evaluate the anti-derivative over the closed interval $\left[0, 1 \right]$ to
find the value of the mean of $X$:

\begin{equation}
\mu = \frac{3}{5}-\frac{18}{16}+\frac{27}{48}\frac{3}{4}.
\end{equation}

\subsubsection{Variance of $X$}
We use the definition of the variance of a continuous random variable in
\cite{reading6a} to compute the variance of $X$.

The definition of Variance Orloff and Bloom give in \cite{reading6a}:

\begin{equation}
  \text{Var}\left(X \right) = E\left( \left( X-\mu\right)^2\right).
\end{equation}

We use the values for $c$ and $\mu$ that we find above to find:
\begin{equation}
  \text{Var}\left(X\right)=int_0^1 x^{2}3\left(x-\frac{3}{4}\right)^2 \,dx.
\end{equation}

Now we multiply some of the factors in the polynomial in the integral above
to get:

\begin{equation}
  \text{Var}\left(X\right)=int_0^1 x^{2}3\left(x^2-\frac{6x}{4}+\frac{9}{16}\right) \,dx.
\end{equation}


We continue multiplying factors:

\begin{equation}
  \text{Var}\left(X\right)=int_0^1 3x^4-\frac{18x^3}{4}+\frac{27x^2}{16} \,dx.
\end{equation}


Now we replace the integral above with its anti-derivative:

\begin{equation}
  \text{Var}\left(X\right)= \frac{3x^4}{5}-\frac{18x^4}{16}+\frac{27x^3}{48} \bigg\rvert_0^1.
\end{equation}

And, we evaluate the anti-derivative over  the interval $\left[0, 1\right]$:

\begin{equation}
  \text{Var}\left(X\right)= \frac{3}{5}-\frac{18}{16}+\frac{27}{48}=\frac{3}{5}-\frac{18}{16}+\frac{9}{16}.
\end{equation}

Now we simplify the expression above further:

\begin{equation}
  \text{Var}\left(X\right)= =\frac{3}{5}-\frac{18}{16}+\frac{9}{16}=\frac{3}{5}-\frac{9}{16}=\frac{48}{80}-\frac{45}{80}=\frac{3}{80}.
\end{equation}

Therefore the variance of $X$ is $\frac{3}{80}$.

\subsubsection{Standard Deviation of $X$}

The standard deviation of $X$ is the square root of its variance \cite{reading6a}.  Therefore the standard deviation of $X$ is:

\begin{equation}
\sigma=\sqrt{\frac{3}{80}} \approx
  0.194.
\end{equation}

\subsubsection{Median value of $X$}
The median value of $X$ is the 0.5
quantile of the cdf of $X$
\cite{reading6a}.

In the first part of this problem,
we find that the pdf of $X$ is
$3x^2$.

Therefore we must solve the equation:
\begin{equation}
int_0^a 3x^2 \,dx = 0.5
\end{equation}

We replace the integral in the
equation above with its
anti-detivative:

\begin{equation}
\frac{3x^3}{3} \bigg\rvert_0^a=0.5.
\end{equation}

And, we evaluate the anti-derivative
above over the interval $\left[0,a
\right]$:

\begin{equation}
\frac{3a^2}{3}=0.5.
\end{equation}

Now it is a matter of doing some
algebra to solve for $a$:

\begin{equation}
3a^2=0.5 \times 3.
\end{equation}

This implies:
\begin{equation}
a^3=0.5.
\end{equation}

Therefore, the median value of $X$ is
$\sqrt[3]{0.5} \approx 0.794$.

\subsection{Standard Deviation of
copies}

For this part of the problem, Orloff
and Bloom give us a set of
random variables $X_1, X_2, \ldots,
X_16$ that are independent, identically-
distributed copies of $X$.
They go on to define $\bar{X}$ as the
average of $X_1, X_2, \ldots, X_16$.

Orloff and Bloom then ask us for the
standard deviation, $\sigma$, of
$\bar{X}$.

$X_1, X_2, \ldots, X_16$ are independent identically-distrubted random
variables.  In \cite{reading6b} Orloff and Bloom show that,for two independent
random variables $X$ and $Y$, $\text{Var}\left( X + Y \right) =
\text{Var}\left( X \right) + \text{Var}\left( Y \right)$.

$\bar{X}$ is the average of $X_1, X_2, \ldots, X_16$.  Therefore

\begin{equation}
\bar{X} = \frac{X_1}{16} + \frac{X_2}{16} + \ldots + \frac{X_16}{16}
\end{equation}

We repeatedly apply the result on the sum of variances of random variables
we quoted from \cite{reading6b} above to get

\begin{equation} \label{varsum1}
\text{Var}\left(\bar{X}\right) = \text{Var}\left(\frac{X_1}{16}\right)
 + \text{Var}\left(\frac{X_2}{16}\right) + \ldots +
 \text{Var}\left(\frac{X_16}{16}\right).
\end{equation}

In \cite{reading6a} Orloff and Bloom state that for
a continuous random variable $Z$,
Var$\left(aZ + b \right)=a^2\text{Var}\left(Z\right)$.

Therefore we may rewrite equation \ref{varsum1} as
\begin{equation} \label{varsum2}
\text{Var}\left(\bar{X}\right) = \frac{1}{16^2}\left(\text{Var}\left(X_1\right)
 + \text{Var}\left(X_2\right) + \ldots +
 \text{Var}\left(X_16\right)\right).
\end{equation}

For this problem, Orloff and Bloom give us that
$X_1 , X_2 , \ldots , X_16$ are identical copies
of $X$, so the variances of $X_1 , X_2 , \ldots , X_16$
are all equal to the variance of $X$.

We computed the variance of $X$ above.
$\text{Var}\left(X \right)=\frac{3}{80}$.

This implies that

\begin{equation}
\text{Var}\left(\bar{X}\right) = \frac{16}{16^2}\left(\frac{3}{80}\right).
\end{equation}

Hence the standard deviation of $\bar{X}$ is
\begin{equation}
\sigma_{\bar{X}} = \sqrt{\frac{3}{80\times16}} \approx 0.0484.
\end{equation}

\subsection{PDF of function of X}
Orloff and Bloom ask us to let $Y=X^4$, and ask us to find the pdf of $Y$.

Let $F_y$ be the cdf of $Y$. Then:

\begin{equation} \label{pofY}
  P\left(Y < y \right) = P \left( X^4 < y \right).
\end{equation}

For $y \neq 0$, equation \ref{pofY} is true if, and only if

\begin{equation} \label{pofY2}
  P\left(Y < y \right) = P \left( X < \sqrt[4]{y} \right).
\end{equation}

The cdf of $X$ is $x^3$. Furthermore, the cdf of $X$ is defined on
the interval $\left[0, 1 \right]$ so equation \ref{pofY2} holds if,
and only if:

\begin{equation}\label{pofY3}
  P\left(Y < y \right) = x^3 \bigg\rvert_0^{\sqrt[4]{y}}.
\end{equation}

We can evaluate equation \ref{pofY3} over the interval
$\left[ 0, \sqrt[4]{y} \right]$; therefore, equation \ref{pofY3} is
true if, and only if:


\begin{equation}\label{pofY3}
  P\left(Y < y \right) = \left( \sqrt[4]{y}\right)^3 .
\end{equation}

Equation \ref{pofY3} is true if, and only if, the cdf of $Y$ is
$\left( \sqrt[4]{y} \right)$.

\section{Gambling}

In this problem Orloff an Bloom give us a
game where the probability
of winning is $0.45$.

Furthermore, they give us two scenarios.
In the first scenario, we bet all the
money we have every time we play.
In the second scenario, we bet \$$5$
each time we play.

Orloff and Bloom ask us which scenario
we should use as a strategty to win
\$$900$, when starting to play with
\$$100$.

We model one round of the game as a
Bernoulli trial with $p=0.45$.

We model the first scenario where
we bet all of the money we have each
time we play with a geometric distribution
where a failure is winning the bet.
The probability of winning \$$1000$
is the probability of four failures before
the first success.

We model the second scenario with a binomial
distribution.

We saw in \cite{reading6b} that the standard normal
approximation of a binomial distributed random variable using
the central limit theorem converges faster
than the standard normal approximatinon an exponentially
distributed random variable.

Therefore the strategy we model with the binomial
distribution approaches the mean value faster than
the model with the exponential distribution.

Since $P=0.45$, the mean value of expected returns
from playing the game is negative.

Therefore we should bet as much as possible
as our strategy to win \$$900$.

\printbibliography{}
\end{document}
